---
title: "Inversion genotyping by PCA"
date: " r format(Sys.Date(),"%B %e, %Y") "
output:
  pdf_document: default
params:
  genotype_file:  ""
  snp_file:  ""
  indiv_file: ""
  log_file:  ""
  evec_file_ann: ""
---

# Inversion genotye checking
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
Genotypes were called from diploid unmasked .fas1k files for the 100kb regions 
around each inversion breakpoint (so 200,000 raw sites). Sites were filtered to 
only sites that have a variant allele, and this was used as the input to smartpca.


## Input files for smartpca
genotype file: `r params$genotype_file` $\\$
snp file: `r params$snp_file` $\\$
individual file: `r params$indiv_file` $\\$

```{r load packages}
suppressPackageStartupMessages( library(ggplot2) )
suppressPackageStartupMessages( library(dplyr) )
suppressPackageStartupMessages( library(knitr) )
suppressPackageStartupMessages( library(kableExtra) )
```

```{r setup input}
input_file   = params$evec_file_ann 
eig = read.table(input_file, header = TRUE)
colnames(eig)[length(colnames(eig))] <- 'INV_STATUS'

eig$INV_STATUS <- factor(eig$INV_STATUS, levels = c("ST", "INV/ST", "INV", "ND"))
levels(eig$INV_STATUS)[4] <-  '<NA>'
#output_file = args[2]
#
```

How many known samples are there in each group?
```{r group counts}
kable(eig %>% count(INV_STATUS)) %>% kableExtra::kable_styling(position = "left")
```

```{r plot PC1 PC2}
p1 <- ggplot(data = eig, aes(x = E1, y = E2, color=INV_STATUS)) + 
	 geom_point(size=2) + 
	 theme_bw() +
	 labs(x = "PC1", y = "PC2")
p1
```

```{r plot PC2 PC3}
p2 <- ggplot(data = eig, aes(x = E2, y = E3, color=INV_STATUS)) + 
	 geom_point(size=2) + 
	 theme_bw() +
	 labs(x = "PC2", y = "PC3")
p2
```

```{r plot PC3 PC4}
p2 <- ggplot(data = eig, aes(x = E3, y = E4, color=INV_STATUS)) + 
	 geom_point(size=2) + 
	 theme_bw() +
	 labs(x = "PC3", y = "PC4")
p2
```

```{r plot PC1 PC2 with geog}
p1 <- ggplot(data = eig, aes(x = E1, y = E2, color=Pop)) + 
	 geom_point(size=2) + 
	 theme_bw() +
	 labs(x = "PC1", y = "PC2")
p1
```

```{r genotyping}
# Call based on PC1 and PC2
d <- as.matrix(dist(eig[,2:3], diag=T, upper=T))

calls <- rep(NA, nrow(eig))

for (i in 1:nrow(eig)) {
  
	top_three <- eig[ sort( d[i,], index.return = TRUE )$ix ,] %>% 
        	        filter(!is.na(INV_STATUS)) %>% select(INV_STATUS) %>% slice(1:3)
	
	# Only want to call if all three agree
	# If all three don't agree, keep call NA
	all_agree <- top_three %>% count(INV_STATUS) %>% filter(n==3) %>% nrow(.)
	
	if (all_agree == 1) {
		top_three %>% count(INV_STATUS) %>% filter(n==3)  %>% select(INV_STATUS) %>% 
		  unlist(., use.names=FALSE) -> calls[i]
	}
}

# need to set as factor
top_three %>% count(INV_STATUS) %>% filter(n==3)  %>% select(INV_STATUS) %>% 
  unlist(., use.names=FALSE) %>% levels(.) -> call_levels
levels(calls) <- call_levels
class(calls) <- "factor"

eig <- cbind(eig, "INV_CALLS"=calls)

kable(eig %>% filter(is.na(INV_STATUS)) %>% select(Library_ID, Pop, INV_CALLS))
```

This plot shows our inversion calls. \
```{r plot calls}
ggplot(data = eig, aes(x = E1, y = E2, color=INV_CALLS)) + 
  geom_point(size=2) + 
  theme_bw() +
  labs(x = "PC1", y = "PC2")
```

Now are there any known samples where this would have failed to call?
```{r check, echo = TRUE}
eig %>% filter(!is.na(INV_STATUS)) %>% filter(INV_STATUS!=INV_CALLS)

```

```{r print indiv, echo = TRUE}
sample_qc_file = paste(strsplit(input_file, '_')[[1]][1], '_mat_sample_qc.tsv', sep='')
print(sample_qc_file)
qc = read.table(sample_qc_file, header = TRUE)
colnames(qc)[1] <- "Library_ID"
```

```{r}
hist(qc$proportion_n)
hist(qc$proportion_het)
```

```{r }
kable(qc)


```

